Clone and graft:\\ Testing scientific applications as they are built

### 2014-12-02 turcksin-spring-mechanics/paper.tex
Write a few more sentences.

%\runningheads{B. Turcksin, T. Heister, W. Bangerth}{Clone and graft: Testing whole applications}

\title{Clone and graft:\\ Testing whole scientific
  applications as they are built}

@@

Second, running and
comparing many tests by hand using the commands above becomes unwieldy. In such
cases, it is useful to write little \texttt{Makefile}s that do this
automatically, or to use one of the tools available as open source -- for
example, the \texttt{CTest} program. It also
becomes more difficult to display tests that create difference when, for
example, you run the 7,000 tests of the \dealii{} testsuite. Again, there are
freely available systems such as \texttt{CDash} that present graphical
overviews of all succeeding and failing tests in a browser.

@@

>>>>

%\runningheads{B. Turcksin, T. Heister, W. Bangerth}{Clone and graft: Testing applications}

\title{Clone and graft:\\ Testing scientific
  applications as they are built}

@@

Second, running and comparing many tests by hand using the commands above
becomes unwieldy. In such cases, it is useful to write little
\texttt{Makefile}s that do this automatically, or to use one of the tools
available as open source -- for example, the \texttt{CTest} program. It also
becomes more difficult to display tests that create differences when, for
example, you run the 7,000 tests of the \dealii{} testsuite. Again, there are
freely available systems such as \texttt{CDash} that present graphical
overviews of all succeeding and failing tests in a browser.

In practice, you're not likely to quickly get to a project size where learning
to use tools like \texttt{CTest} and \texttt{CDash} are important
considerations. In practice, the point where using such tools becomes
indispensible is when your code has several 10,000 lines of code or more --
several years worth of work. When you get to this point, your project will
have all sorts of other growing pains as well, such as having to deal with
external users, running mailing lists, portability, managing complexity and
other issues. Some of these are discussed in \cite{BH13}. The point simply is
that there is no need to worry too much about testing tools for programs of
this size: care about writing tests from the beginning and if the number of
tests becomes too unwieldy, there will be professional-grade tools you can
switch to that can manage this for you. In the meantime, a little shell script
such as the following will do just fine:
\begin{lstlisting}[frame=single,basicstyle=\footnotesize,language=bash]
  # loop over all .json input files
  for jsonfile in *.json ; do
    # execute our program on it and pipe the output into a file
    python ${jsonfile} > ${jsonfile}.new

    # then compare output; only ask 'diff' to report whether the
    # files are the same or not, and create output depending on this:
    # print an 'X' if there is a difference, a '.' if there isn't. this
    # ensures that test failures are easily visible
    if (diff ${jasonfile} ${jsonfile}.new > /dev/null) ; then
      echo " .    ${jasonfile}" ;
    else 
      echo " X    ${jasonfile}" ;
    fi
  done
\end{lstlisting}
If you call this script, say, \texttt{tests/compare.sh}, then running it using
the command \texttt{cd tests/ ; bash -e compare.sh} should produce output like
this:
\begin{lstlisting}[frame=single,basicstyle=\footnotesize,language=bash]
 .    testcase-1.json
 .    testcase-2.json
 .    testcase-3.json
 .    testcase-4.json
 .    testcase-5.json
\end{lstlisting}
On the other hand, if your output looks like this
\begin{lstlisting}[frame=single,basicstyle=\footnotesize,language=bash]
 .    testcase-1.json
 .    testcase-2.json
 .    testcase-3.json
 X    testcase-4.json
 .    testcase-5.json
\end{lstlisting}
then you know that you just broke some functionality that
\texttt{testcase-4.json} checks. Given how easy it is to run this script, you
want to run this script often to find new bugs as early as possible.

@@

Scientific computing deals, at a fundamental level, almost always with
floating point numbers. Unfortunately, floating point arithmetic suffers from
some rather annoying oddities that can make testing scientific codes more
challenging than code that doesn't use floating point numbers.

First, floating point numbers do not enjoy the same properties as integers.
